LCS Tool : A Computational Platform for Lagrangian Coherent Structures

### 2014-08-23 main.tex
Make style consistent in tables
Change lambda-line LCSs to elliptic LCSs

\begin{table}
\begin{center}
\begin{tabular}{|c|p{.7\textwidth}|}
\hline
\multicolumn{2}{|p{\textwidth}|}{\lstinline![cgEigenvector,cgEigenvalue] = eig_cgStrain(derivative,domain,timespan,resolution)!}\tabularnewline
\hline
\lstinline!derivative! & function handle for flow velocity equations\tabularnewline
\hline
\lstinline!domain! & $2 \times 2$ array to define flow domain\tabularnewline
\hline
\lstinline!timespan! & $1 \times 2$ array to define flow timespan\tabularnewline
\hline
\lstinline!resolution! & $1 \times 2$ array to define Cauchy-Green strain main grid resolution\tabularnewline
\hline
\lstinline!auxGridRelDelta! & optional scalar between 0 and 0.5 to specify auxiliary grid spacing. Default is $10^{-2}$.\tabularnewline
\hline
\lstinline!eigenvalueFromMainGrid! & optional logical to control whether eigenvalues of Cauchy-Green strain are calculated from main or auxiliary grid. Default is \lstinline!true!.\tabularnewline
\hline
\lstinline!incompressible! & optional logical to specify if incompressibility is imposed. Default is \lstinline!false!.\tabularnewline
\hline
\lstinline!odeSolverOptions! & optional \lstinline!odeset! structure to specify flow map integration parameters\tabularnewline
\hline
\end{tabular}
\caption{Syntax of the function \lstinline!eig_cgStrain!}
\label{t:eig_cgStrain syntax}
\end{center}
\end{table}

@@

\begin{table}
\begin{tabular}{|c|p{.7\textwidth}|}
\hline
\multicolumn{2}{|p{\textwidth}|}
{\lstinline![shearline.etaPos,shearline.etaNeg] = lambda_line(cgEigenvector,cgEigenvalue,lambda)!}\tabularnewline
\hline
\lstinline!cgEigenvector! & array of Cauchy-Green strain eigenvectors\tabularnewline
\hline
\lstinline!cgEigenvalue! & array of Cauchy-Green strain eigenvalues\tabularnewline
\hline
\lstinline!lambda! & scalar lambda value in \cref{eq:eta}\tabularnewline
\hline \hline
\multicolumn{2}{|p{\textwidth}|}{\lstinline![closedOrbits,orbits] = poincare_closed_orbit_multi(domain,resolution,shearline,PSList)!}\tabularnewline
\hline
\lstinline!domain! & array to define flow domain\tabularnewline
\hline
\lstinline!resolution! & $1 \times 2$ array to define main grid resolution for Cauchy-Green strain tensor\tabularnewline
\hline
\lstinline!shearline! & structure of arrays of $\eta_+$ and $\eta_-$ values on main grid\tabularnewline
\hline
\lstinline!PSList! & user-defined structure for Poincare section end-points, number of $\lambda$-lines launched from Poincare section, and maximum closed $\lambda$-line length\tabularnewline
\hline
\lstinline!nBisection! & optional number of bisection steps to refine zero crossings of Poincare map. Default value: 5.\tabularnewline
\hline
\lstinline!dThresh! & optional threshold to discard discontinuous zero crossings of Poincare map. Default value: $10^{-2}$.\tabularnewline
\hline
\lstinline!odeSolverOptions! & optional \lstinline!odeset! structure to specify $\lambda$-line integration parameters\tabularnewline
\hline
\lstinline!periodicBc! & optional $1 \times 2$ logical array to specify periodic boundary conditions. Default is \lstinline![false,false]!.\tabularnewline
\hline
\end{tabular}
\caption{Syntax of LCS Tool's elliptic LCS functions.}
\label{t:Elliptic LCS functions}
\end{table}

@@

\begin{table}
\begin{center}
\begin{tabular}{|c|p{.7\textwidth}|}
\hline
\multicolumn{2}{|p{\textwidth}|}
{\lstinline![curvePosition,curveInitialPosition] = seed_curves_from_lambda_max(distance,cgEigenvalue,cgEigenvector,flowDomain,flowResolution)!}\tabularnewline
\hline
\lstinline!distance! & threshold distance for placement of $\lambda_2(x_0)$ maxima\tabularnewline
\hline
\lstinline!cgEigenvalue! & array of Cauchy-Green strain eigenvalues\tabularnewline
\hline
\lstinline!cgEigenvector! & array of Cauchy-Green strain eigenvectors\tabularnewline
\hline
\lstinline!flowDomain! & $2 \times 2$ array to define flow domain\tabularnewline
\hline
\lstinline!flowResolution! & $1 \times 2$ array to define Cauchy-Green strain main grid resolution\tabularnewline
\hline
\lstinline!periodicBc! & optional $1 \times 2$ logical array to specify periodic boundary conditions. Default is \lstinline![false,false]!.\tabularnewline
\hline
\lstinline!nMaxCurves! & optional maximum number of curves (i.e. shrinklines of stretchlines) to generate. Default is \lstinline!numel(cgEigenvalue)!.\tabularnewline
\hline
\lstinline!odeSolverOptions! & optional \lstinline!odeset! structure to specify flow map integration parameters\tabularnewline
\hline
\end{tabular}
\end{center}
\caption{Syntax of the function \lstinline!seed_curves_from_lambda_max!}
\label{t:seed_curves_from_lambda_max syntax}
\end{table}

@@

\lstinputlisting[caption={Double gyre script for elliptic lambda-line LCSs},label=l:double gyre lambda-line LCS]{listings/double_gyre/lambda_lcs.m}

@@

\begin{figure}
\centering
\includegraphics[width=\textwidth]{graphics/double_gyre/lambda_strain_lcs}
\caption{Double gyre lambda-line and shrinkline LCSs}
\label{fig:double_gyre_lambda_strain_lcs}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=\textwidth]{graphics/double_gyre/lambda_stretch_lcs}
\caption{Double gyre lambda-line and stretchline LCSs}
\label{fig:double_gyre_lambda_stretch_lcs}
\end{figure}

@@

The integration time is $T=4L_x/U$, a multiple of a the eddy turnover
time in the flow. \Cref{fig:Bickley_jet_lambda_LCS_full} shows lambda-line LCS of the Bickley jet with the FTLE in the background. The figure demonstrates a poor result since the LCSs change shape significantly between eddies. To address this problem we could adjust numerical parameters such as the resolution, but instead we will analyze lambda-line closed orbits of one vortex for a short timespan.

\begin{figure}
\centering
\includegraphics[width=.85\textwidth]{graphics/bickley_jet/lambda_lcs_full}
\caption{Bickley jet lambda-line LCSs}
\label{fig:Bickley_jet_lambda_LCS_full}
\end{figure}

\Cref{fig:Bickley jet lambda lcs convergence} shows convergence of the lambda-line elliptic LCSs. The LCS Tool script used to produce that figure are based on the script \lstinline!demo/bickley_jet/hyperbolic_shear_lcs.m! which is given in \cref{lst:Bickley jet hyperbolic_shear_lcs subset}. It is worth pointing out that in order to obtain closed orbits, $\lambda = 0.995$; with $\lambda = 1$ no closed orbits exist.

@@

\begin{figure}[hbt]
\centering
\includegraphics[width=.7\textwidth]{graphics/bickley_jet/lambda_lcs_convergence_400}
\includegraphics[width=.7\textwidth]{graphics/bickley_jet/lambda_lcs_convergence_500}
\includegraphics[width=.7\textwidth]{graphics/bickley_jet/lambda_lcs_convergence_600}
\caption{Bickley jet closed lambda-line at varying resolutions}
\label{fig:Bickley jet lambda lcs convergence}
\end{figure}

@@

% Lambda-line LCSs
% Define Poincare sections; ...
poincareSection = struct('endPosition',{},'numPoints',{},'orbitMaxLength',{});
...
poincareSection(1).endPosition = [3.3,-32.1;3.7,-31.6];
poincareSection(2).endPosition = [1.3,-30.9;1.9,-31.1];

>>>>

\begin{table}
\begin{center}
\begin{tabular}{|c|p{.7\textwidth}|}
\hline
\multicolumn{2}{|p{\textwidth}|}{\lstinline![cgEigenvector,cgEigenvalue] = eig_cgStrain(derivative,domain,timespan,resolution)!}\tabularnewline
\hline
\lstinline!derivative! & function handle for flow velocity equations\tabularnewline
\hline
\lstinline!domain! & $2 \times 2$ array to define flow domain\tabularnewline
\hline
\lstinline!timespan! & $1 \times 2$ array to define flow timespan\tabularnewline
\hline
\lstinline!resolution! & $1 \times 2$ array to define Cauchy-Green strain main grid resolution\tabularnewline
\hline
\lstinline!auxGridRelDelta! & optional scalar between 0 and 0.5 to specify auxiliary grid spacing. Default: $10^{-2}$.\tabularnewline
\hline
\lstinline!eigenvalueFromMainGrid! & optional logical to control whether eigenvalues of Cauchy-Green strain are calculated from main or auxiliary grid. Default: \lstinline!true!.\tabularnewline
\hline
\lstinline!incompressible! & optional logical to specify if incompressibility is imposed. Default: \lstinline!false!.\tabularnewline
\hline
\lstinline!odeSolverOptions! & optional \lstinline!odeset! structure to specify flow map integration parameters\tabularnewline
\hline
\end{tabular}
\caption{Syntax of the function \lstinline!eig_cgStrain!}
\label{t:eig_cgStrain syntax}
\end{center}
\end{table}

@@

\begin{table}
\begin{tabular}{|c|p{.7\textwidth}|}
\hline
\multicolumn{2}{|p{\textwidth}|}
{\lstinline![shearline.etaPos,shearline.etaNeg] = lambda_line(cgEigenvector,cgEigenvalue,lambda)!}\tabularnewline
\hline
\lstinline!cgEigenvector! & array of Cauchy-Green strain eigenvectors\tabularnewline
\hline
\lstinline!cgEigenvalue! & array of Cauchy-Green strain eigenvalues\tabularnewline
\hline
\lstinline!lambda! & scalar lambda value in \cref{eq:eta}\tabularnewline
\hline \hline
\multicolumn{2}{|p{\textwidth}|}{\lstinline![closedOrbits,orbits] = poincare_closed_orbit_multi(domain,resolution,shearline,PSList)!}\tabularnewline
\hline
\lstinline!domain! & array to define flow domain\tabularnewline
\hline
\lstinline!resolution! & $1 \times 2$ array to define main grid resolution for Cauchy-Green strain tensor\tabularnewline
\hline
\lstinline!shearline! & structure of arrays of $\eta_+$ and $\eta_-$ values on main grid\tabularnewline
\hline
\lstinline!PSList! & user-defined structure for Poincare section end-points, number of $\lambda$-lines launched from Poincare section, and maximum closed $\lambda$-line length\tabularnewline
\hline
\lstinline!nBisection! & optional number of bisection steps to refine zero crossings of Poincare map. Default: 5.\tabularnewline
\hline
\lstinline!dThresh! & optional threshold to discard discontinuous zero crossings of Poincare map. Default: $10^{-2}$.\tabularnewline
\hline
\lstinline!odeSolverOptions! & optional \lstinline!odeset! structure to specify $\lambda$-line integration parameters\tabularnewline
\hline
\lstinline!periodicBc! & optional $1 \times 2$ logical array to specify periodic boundary conditions. Default: \lstinline![false,false]!.\tabularnewline
\hline
\end{tabular}
\caption{Syntax of LCS Tool's elliptic LCS functions.}
\label{t:Elliptic LCS functions}
\end{table}

@@

\begin{table}
\begin{center}
\begin{tabular}{|c|p{.7\textwidth}|}
\hline
\multicolumn{2}{|p{\textwidth}|}
{\lstinline![curvePosition,curveInitialPosition] = seed_curves_from_lambda_max(distance,cgEigenvalue,cgEigenvector,flowDomain,flowResolution)!}\tabularnewline
\hline
\lstinline!distance! & threshold distance for placement of $\lambda_2(x_0)$ maxima\tabularnewline
\hline
\lstinline!cgEigenvalue! & array of Cauchy-Green strain eigenvalues\tabularnewline
\hline
\lstinline!cgEigenvector! & array of Cauchy-Green strain eigenvectors\tabularnewline
\hline
\lstinline!flowDomain! & $2 \times 2$ array to define flow domain\tabularnewline
\hline
\lstinline!flowResolution! & $1 \times 2$ array to define Cauchy-Green strain main grid resolution\tabularnewline
\hline
\lstinline!periodicBc! & optional $1 \times 2$ logical array to specify periodic boundary conditions. Default: \lstinline![false,false]!.\tabularnewline
\hline
\lstinline!nMaxCurves! & optional maximum number of curves (i.e. shrinklines of stretchlines) to generate. Default: \lstinline!numel(cgEigenvalue)!.\tabularnewline
\hline
\lstinline!odeSolverOptions! & optional \lstinline!odeset! structure to specify flow map integration parameters\tabularnewline
\hline
\end{tabular}
\end{center}
\caption{Syntax of the function \lstinline!seed_curves_from_lambda_max!}
\label{t:seed_curves_from_lambda_max syntax}
\end{table}

@@

\lstinputlisting[caption={Double gyre script for elliptic LCSs},label=l:double gyre lambda-line LCS]{listings/double_gyre/lambda_lcs.m}

@@

\begin{figure}
\centering
\includegraphics[width=\textwidth]{graphics/double_gyre/lambda_strain_lcs}
\caption{Double gyre elliptic and shrinkline LCSs}
\label{fig:double_gyre_lambda_strain_lcs}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=\textwidth]{graphics/double_gyre/lambda_stretch_lcs}
\caption{Double gyre elliptic and stretchline LCSs}
\label{fig:double_gyre_lambda_stretch_lcs}
\end{figure}

@@

The integration time is $T=4L_x/U$, a multiple of a the eddy turnover
time in the flow. \Cref{fig:Bickley_jet_lambda_LCS_full} shows elliptic LCSs of the Bickley jet with the FTLE in the background. The figure demonstrates a poor result since the LCSs change shape significantly between eddies. To address this problem we could adjust numerical parameters such as the resolution, but instead we will analyze lambda-line closed orbits of one vortex for a short timespan.

\begin{figure}
\centering
\includegraphics[width=.85\textwidth]{graphics/bickley_jet/lambda_lcs_full}
\caption{Bickley jet elliptic LCSs}
\label{fig:Bickley_jet_lambda_LCS_full}
\end{figure}

\Cref{fig:Bickley jet lambda lcs convergence} shows convergence of the elliptic LCSs. The LCS Tool script used to produce that figure are based on the script \lstinline!demo/bickley_jet/hyperbolic_shear_lcs.m! which is given in \cref{lst:Bickley jet hyperbolic_shear_lcs subset}. It is worth pointing out that in order to obtain closed orbits, $\lambda = 0.995$; with $\lambda = 1$ no closed orbits exist.

@@

\begin{figure}[hbt]
\centering
\includegraphics[width=.7\textwidth]{graphics/bickley_jet/lambda_lcs_convergence_400}
\includegraphics[width=.7\textwidth]{graphics/bickley_jet/lambda_lcs_convergence_500}
\includegraphics[width=.7\textwidth]{graphics/bickley_jet/lambda_lcs_convergence_600}
\caption{Bickley jet closed lambda-lines at varying resolutions}
\label{fig:Bickley jet lambda lcs convergence}
\end{figure}

@@

% Elliptic LCSs
% Define Poincare sections; ...
poincareSection = struct('endPosition',{},'numPoints',{},'orbitMaxLength',{});
...
poincareSection(1).endPosition = [3.3,-32.1;3.7,-31.6];
poincareSection(2).endPosition = [1.3,-30.9;1.9,-31.1];
